---
title: "CEMA 0907: Statistics in the Real World"
subtitle: "Extra Topics"
author: "Anthony Scotina"
date: 
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: my-theme.css
    nature:
      countIncrementalSlides: false
      highlightLines: true
---

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_accent(base_color = "#43418A")
```

# Needed Packages

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidytext) # Install this!
```

---

class: center, middle, frame

# Sentiment Analysis

## (Anthony's Version)

---

# Sentiment Analysis

Using *text mining techniques*, one piece of information we can extract from text is the **sentiment of the text**. 

- Is the text positive or negative? Does it evoke *surprise* or *disgust*?

- If we *sum* the sentiment for *each word in a document*, does this reflect an accurate measure of sentiment for the document as a whole?

--

There are a variety of ways to judge the **sentiment** of a text. 

- One such *sentiment lexicon* is the `bing` lexicon from [Bing Liu and collaborators](https://www.cs.uic.edu/~liub/FBS/sentiment-analysis.html). 

From `tidytext`:

```{r, eval = FALSE}
get_sentiments("bing")
```

---

# Taylor Swift

I would **Love** to tell a **Story** about the sentiment of Taylor Swift's songs. 

From [TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-09-29/readme.md) via Rosie Baillie and Dr. Sara Stoudt:

```{r}
taylor_swift_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/taylor_swift_lyrics.csv')
```

---

# Tidy Text

Let's use `unnest_tokens()` to convert the data to *tidy* format!

- Each word will appear in its own row.

```{r}
tswift_tidy = taylor_swift_lyrics %>%
  unnest_tokens(output = word, input = Lyrics)
```

--

First, we'll want to *remove* **stop words**:

```{r}
head(stop_words) # 'stop_words' from 'tidytext' package
```

---

# `anti_join()`

Here's where we can use `anti_join()`!

- *Remove* stop words from `tswift_tidy`. 

```{r}
tswift_tidy_anti = tswift_tidy %>%
  anti_join(stop_words, by = "word")
```

---

# Which words does T. Swift say the most?

```{r, dpi = 300}
tswift_tidy_anti %>%
  count(word, sort = TRUE) %>% 
  slice_max(n, n = 22) %>%
  ggplot(aes(x = fct_reorder(word, n), y = n)) + 
  geom_col() + 
  labs(x = "", y = "Count", 
       title = "Which words does Taylor Swift LOVE the most?") +
  coord_flip() + 
  theme_minimal()
```

---

# Back to Sentiment Analysis

The text is in **tidy format** (one word per row), so we can proceed with the **sentiment analysis**. 

Let's **join** the `bing` lexicon to `tswift_tidy_anti`. 

- We'll use `inner_join()`, because we want only the words that appear in **both** the song and the `bing` lexicon:

```{r}
song_sentiment = tswift_tidy_anti %>%
  inner_join(get_sentiments("bing"), by = "word")
```

---

# Taylor Swift Sentiment Analysis

Next, we'll `count()` up the `word` and `sentiment` combinations, extract the **top 22** for each sentiment, and pass them off to `ggplot()` and `geom_col()`. 

- I used `scales = "free_y"` in `facet_wrap()` so that each facet contained a different set of words (positive vs negative sentiments).

```{r, eval = FALSE}
sentiment_counts = song_sentiment %>%
  count(word, sentiment, sort = TRUE)

sentiment_counts %>%
  group_by(sentiment) %>%
  slice_max(order_by = n, n = 10) %>%
  ggplot(aes(x = fct_reorder(word, n), y = n, 
             fill = sentiment)) + 
  geom_col() + 
  facet_wrap( ~ sentiment, scales = "free_y") + 
  labs(x = "", y = "Contribution to sentiment", 
       title = "Sentiment Analysis of Taylor Swift Songs") +
  theme_bw() + 
  theme(legend.position = "none") +
  coord_flip()
```

---

# Taylor Swift Sentiment Analysis

.center[
```{r, echo = FALSE, dpi = 300, out.width = "65%"}
sentiment_counts = song_sentiment %>%
  count(word, sentiment, sort = TRUE)

sentiment_counts %>%
  group_by(sentiment) %>%
  top_n(10, wt = n) %>%
  ggplot(aes(x = fct_reorder(word, n), y = n, 
             fill = sentiment)) + 
  geom_col() + 
  facet_wrap( ~ sentiment, scales = "free_y") + 
  labs(x = "", y = "Contribution to sentiment", 
       title = "Sentiment Analysis of Taylor Swift Songs") +
  theme_bw() + 
  theme(legend.position = "none") +
  coord_flip()
```
]

---

class: center, middle

# One More Thing

## case_when()

---

# `case_when()`

Are there more **resort** hotel reservations (compared to **city** hotel reservations) during the *summer* months?

This dataset contains open data on **hotel booking demand** via [TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md). 

```{r, warning = FALSE, message = FALSE}
hotels <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv')
```

--

**Two variables of interest**:

- `hotel`: `city` or `resort`
- `arrival_date_month`: the *month* (in words) of arrival

---

# `case_when()`

From `?case_when`:

> This function allows you to vectorise multiple `if_else()` statements. 

--

Let's create a new `season` variable:

- If `arrival_date_month %in% c("December", "January", "February")`, then `season = "Winter"`

- If `arrival_date_month %in% c("March", "April", "May")`, then `season = "Spring"`

- If `arrival_date_month %in% c("June", "July", "August")`, then `season = "Summer"`

- If `arrival_date_month %in% c("September", "October", "November")`, then `season = "Fall"`

---

# `case_when()`

From `?case_when`:

> This function allows you to vectorise multiple `if_else()` statements. 

The general syntax within a `case_when()` statement is: 

.center[
case_when(`CONDITION1 ~ CATEGORY NAME IF CONDITION1 IS TRUE`, 
`CONDITION1 ~ CATEGORY NAME IF CONDITION1 IS TRUE`
]
```{r, eval = FALSE}
mutate(new_variable = 
         case_when(`CONDITION1 ~ CATEGORY NAME IF CONDITION1 IS TRUE`, 
                   `CONDITION2 ~ CATEGORY NAME IF CONDITION2 IS TRUE`, 
                   ... )
       )
```

---

# `case_when()`

From `?case_when`:

> This function allows you to vectorise multiple `if_else()` statements. 

Let's create a new `season` variable:

```{r}
hotels = hotels %>%
  mutate(season = 
           case_when(arrival_date_month %in% c("December", "January", "February") ~ "Winter", 
                     arrival_date_month %in% c("March", "April", "May") ~ "Spring", 
                     arrival_date_month %in% c("June", "July", "August") ~ "Summer", 
                     arrival_date_month %in% c("September", "October", "November") ~ "Fall"))

```

---

# Hotel Reservations (by season)

Now let's make a bar graph using the new `season` variable!

```{r, eval = FALSE}
hotels %>%
  ggplot(aes(x = season, fill = hotel)) + 
  geom_bar(position = "dodge") + 
  labs(x = "", y = "", fill = "",
       title = "Hotel Reservations (by season)") + 
  scale_y_continuous(labels = scales::comma) +
  coord_flip() +
  theme_minimal()
```